---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

### Loading and preprocessing the data
```{r}
data <- read.csv( "activity.csv" )
data$date <- as.POSIXlt(data$date)
data$wdayname <- weekdays(as.Date(data$date))
```

### What is mean total number of steps taken per day?
1. Calculate the total number of steps taken per day
2. Calculate and report the mean and median of the total number of steps taken per day
```{r}
totalperday <- tapply( data$step, data$wdayname, sum, na.rm=TRUE )
meanperday <- tapply( data$step, data$wdayname, mean, na.rm=TRUE )
medianperday <- tapply( data$step, data$wdayname, median, na.rm=TRUE)

statsperday <- data.frame( total=totalperday, mean=meanperday, median=medianperday )
statsperday
```
3. Make a histogram of the total number of steps taken each day
```{r}
notNAData <- data[!is.na(data$step), "date"]
hist( notNAData$wday, col="grey" )
```

### What is the average daily activity pattern?
1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
meanday <- tapply( data$step, data$interval, mean, na.rm=TRUE )
plot( meanday, type="l" )
```


### Imputing missing values
1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
sum( is.na(data$step) ) # Total number of missing observations.
mean( is.na(data$step) )*100 # Percentage of missing observations.
```
2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

Answer: **Inteval mean will be used since it is relatively easy to generte and will not bias the data.**

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
dataNNA <- data
intervalmean <- tapply( dataNNA$steps, dataNNA$interval, mean ) # Mean by interval

# Replace each step == NA with the interval mean
for( i in dataNNA$interval )
{
    dataNNA$steps[dataNNA$interval==i] <- 
        as.numeric( intervalmean[as.character(i)] )    
}
```

4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

Answer: ** The frequence of the observations per day was altered. The histogram
shows differences in Tuesday, Thursday, Friday and Saturday.**
```{r}
notNAData <- dataNNA[, "date"]
hist( notNAData$wday, col="grey" )
```